---
title: "amcModels-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{amcModels-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



Throughout this vignette we will demonstrate the uses of the `amcModels` package on a simple example using the following transition matrices across 6 groups with 3 transitory tactics and 2 absorbing states (5 total states).

```{r}
P1 <- P2 <- matrix(c(.3, .2, .25, .15, .1, 
                     .1, .3, .4, .1, .1,
                     .05, .5, .2, .1, .15), nrow = 3, ncol = 5,
                   byrow = TRUE)
P3 <- matrix(c(.1, .3, .3, .15, .15, 
               .1, .3, .4, .1, .1,
               .5, .3, .1, .05, .05), nrow = 3, ncol = 5,
              byrow = TRUE)
P4 <- matrix(c(.15, .25, .3, .2, .1, 
               .15, .2, .45, .1, .1,
               .4, .35, .1, .1, .05), nrow = 3, ncol = 5,
              byrow = TRUE)
P5 <- matrix(c(.1, .3, .3, .15, .15, 
               .3, .3, .25, .1, .05,
               .5, .3, .1, .05, .05), nrow = 3, ncol = 5,
              byrow = TRUE)
P6 <- matrix(c(.6, .1, .1, .1, .1, 
               .3, .4, .1, .1, .1,
               .05, .3, .45, .05, .15), nrow = 3, ncol = 5,
              byrow = TRUE)
```


```{r setup}
library(amcModels)
```


### Step1: Simulate sequences:

For each group we will simulate 100 sequences of events starting in each state beginning in each of the three transitory states for a total of 300 sequences.

```{r}
n <- 100

## Simulate data starting in state 1:
data_1_1 <- absmarkovchain(1, P1, 100)
data_2_1 <- absmarkovchain(1, P2, 100)
data_3_1 <- absmarkovchain(1, P3, 100)
data_4_1 <- absmarkovchain(1, P4, 100)
data_5_1 <- absmarkovchain(1, P5, 100)
data_6_1 <- absmarkovchain(1, P6, 100)

## Simulate data starting in state 2:
data_1_2 <- absmarkovchain(2, P1, 100)
data_2_2 <- absmarkovchain(2, P2, 100)
data_3_2 <- absmarkovchain(2, P3, 100)
data_4_2 <- absmarkovchain(2, P4, 100)
data_5_2 <- absmarkovchain(2, P5, 100)
data_6_2 <- absmarkovchain(2, P6, 100)

## Simulate data starting in state 3:
data_1_3 <- absmarkovchain(3, P1, 100)
data_2_3 <- absmarkovchain(3, P2, 100)
data_3_3 <- absmarkovchain(3, P3, 100)
data_4_3 <- absmarkovchain(3, P4, 100)
data_5_3 <- absmarkovchain(3, P5, 100)
data_6_3 <- absmarkovchain(3, P6, 100)
```

### Step 2: Create data array

```{r}
N <- array(0, dim = c(6, 3, 5))

## Function to apply to each sequence:
make_counts <- function(seq, N_group){
  for(x in 2:length(seq)){
    row.ref <- seq[x - 1]
    col.ref <- seq[x]
    N_group[row.ref, col.ref] <- N_group[row.ref, col.ref] + 1
  }
  return(N_group)
}

for(i in 1:100){
  ## Group 1:
  N[1, , ] <- make_counts(data_1_1[[i]], N[1, , ])
  N[1, , ] <- make_counts(data_1_2[[i]], N[1, , ])
  N[1, , ] <- make_counts(data_1_3[[i]], N[1, , ])
  
  ## Group 2:
  N[2, , ] <- make_counts(data_2_1[[i]], N[2, , ])
  N[2, , ] <- make_counts(data_2_2[[i]], N[2, , ])
  N[2, , ] <- make_counts(data_2_3[[i]], N[2, , ])
  
  ## Group 3:
  N[3, , ] <- make_counts(data_3_1[[i]], N[3, , ])
  N[3, , ] <- make_counts(data_3_2[[i]], N[3, , ])
  N[3, , ] <- make_counts(data_3_3[[i]], N[3, , ])
  
  ## Group 4:
  N[4, , ] <- make_counts(data_4_1[[i]], N[4, , ])
  N[4, , ] <- make_counts(data_4_2[[i]], N[4, , ])
  N[4, , ] <- make_counts(data_4_3[[i]], N[4, , ])
  
  ## Group 5:
  N[5, , ] <- make_counts(data_5_1[[i]], N[5, , ])
  N[5, , ] <- make_counts(data_5_2[[i]], N[5, , ])
  N[5, , ] <- make_counts(data_5_3[[i]], N[5, , ])
  
  ## Group 6:
  N[6, , ] <- make_counts(data_6_1[[i]], N[6, , ])
  N[6, , ] <- make_counts(data_6_2[[i]], N[6, , ])
  N[6, , ] <- make_counts(data_6_3[[i]], N[6, , ])
}
```

## Step 4: Fit models

First we fit the AMC model which will return posterior draws for a combined transition matrix.

```{r}
fit_amc <- amc(N, xsi = 1/5)
```

Second we will fit the HAMC model.

```{r}
fit_hamc <- hamc(N, MCMC.cores = 3)
```

Lastly we will fit the EPA-HAMC model with the sequential updating of clusters and thinning by 5.

```{r}
fit_epa <- epa_hamc(N, method = "seq", thin = 5, MCMC.cores = 3)
```

## Step 5: Analyze results

We we explore how we would find answers to the following questions:

- Which transitory state (1, 2, or 3) is most likely to lead to either of the absorbing states (4 or 5)?
- Which of the six groups are clustered most often together across all three transitory states?
- Which of the six groups differs the most from the average?

First lets suppose are main interest is which transitory state (1, 2, or 3) is the most likely to lead to either of the absorbing states we can examine the posterior densities of the transition matrices. We can use any of the three fited models to analyze this. First we will explore the results from the AMC model.

```{r}
## Look at posterior densities of going into an absorbing state:
# Make density plot of going into state 4:
plot(density(fit_amc[ , 1, 4]), col = "red",
     main = "Probability of going into state 4",
     xlim = c(0, .2),
     ylim = c(0, 100), xlab = "")
lines(density(fit_amc[ , 2, 4]), col = "blue")
lines(density(fit_amc[ , 3, 4]), col = "black")
legend("topleft", legend = c("1", "2", "3"), col = c("red", "blue", "black"),
       lty = 1)

## Make density plot of going into state 5:
plot(density(fit_amc[ , 1, 5]), col = "red",
     main = "Probability of going into state 5",
     xlim = c(0, .2),
     ylim = c(0, 100))
lines(density(fit_amc[ , 2, 5]), col = "blue")
lines(density(fit_amc[ , 3, 5]), col = "black")
legend("top", legend = c("1", "2", "3"), col = c("red", "blue", "black"),
       lty = 1, horiz = TRUE)

```

Analyzing this we see that it shows that state 1 is most likely to lead to either of the absorbing states.
