---
title: "amcModels-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{amcModels-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Throughout this vignette we will demonstrate the uses of the `amcModels` package on a simple example using the following transition matrices across 6 groups with 3 transitory tactics and 2 absorbing states (5 total states).

```{r}
P1 <- P2 <- matrix(c(.3, .2, .25, .15, .1, 
                     .1, .3, .4, .1, .1,
                     .05, .5, .2, .1, .15), nrow = 3, ncol = 5,
                   byrow = TRUE)
P3 <- matrix(c(.1, .3, .3, .15, .15, 
               .1, .3, .4, .1, .1,
               .5, .3, .1, .05, .05), nrow = 3, ncol = 5,
              byrow = TRUE)
P4 <- matrix(c(.15, .25, .3, .2, .1, 
               .15, .2, .45, .1, .1,
               .4, .35, .1, .1, .05), nrow = 3, ncol = 5,
              byrow = TRUE)
P5 <- matrix(c(.1, .3, .3, .15, .15, 
               .3, .3, .25, .1, .05,
               .5, .3, .1, .05, .05), nrow = 3, ncol = 5,
              byrow = TRUE)
P6 <- matrix(c(.6, .1, .1, .1, .1, 
               .3, .4, .1, .1, .1,
               .05, .3, .45, .05, .15), nrow = 3, ncol = 5,
              byrow = TRUE)
```


```{r setup}
library(amcModels)
```


### Step1: Simulate sequences:

For each group we will simulate 100 sequences of events starting in each state beginning in each of the three transitory states for a total of 300 sequences.

```{r}
n <- 100

## Simulate data starting in state 1:
data_1_1 <- absmarkovchain(1, P1, 100)
data_2_1 <- absmarkovchain(1, P2, 100)
data_3_1 <- absmarkovchain(1, P3, 100)
data_4_1 <- absmarkovchain(1, P4, 100)
data_5_1 <- absmarkovchain(1, P5, 100)
data_6_1 <- absmarkovchain(1, P6, 100)

## Simulate data starting in state 2:
data_1_2 <- absmarkovchain(2, P1, 100)
data_2_2 <- absmarkovchain(2, P2, 100)
data_3_2 <- absmarkovchain(2, P3, 100)
data_4_2 <- absmarkovchain(2, P4, 100)
data_5_2 <- absmarkovchain(2, P5, 100)
data_6_2 <- absmarkovchain(2, P6, 100)

## Simulate data starting in state 3:
data_1_3 <- absmarkovchain(3, P1, 100)
data_2_3 <- absmarkovchain(3, P2, 100)
data_3_3 <- absmarkovchain(3, P3, 100)
data_4_3 <- absmarkovchain(3, P4, 100)
data_5_3 <- absmarkovchain(3, P5, 100)
data_6_3 <- absmarkovchain(3, P6, 100)
```

### Step 2: Create data array

```{r}
N <- array(NA, dim = c(6, 3, 5))
```

